Alleviating the Topology Mismatch Problem in Unstructured and Structured Overlay Networks: A Survey

### 2013-11-10 distrproj.tex
16.03.2010 (Vassilis) Restructuring initial approach.

%opening
\title{Alleviating the Topology Mismatch Problem with Location-Aware Overlays: A Survey}
\author{
Vassilis Moustakas\\
Department of Informatics and Telecommunications,\\
National and Kapodistrian University of Athens\\
\mbox{\texttt{b.moystakas@di.uoa.gr}}
}

\begin{document}

\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
Peer-to-peer systems is a rapidly growing application architecture. The overlay network abstraction on top of a best-effort infrastructure provides a powerfull tool to realize ever-wanted exotic features like anonymity, robustness, load balancing, QoS at the application layer and many more. Unfortunately weaknesses of early effectuations constrained applications from unleashing the full potential of the paradigm. One of them is the topology mismatch problem between the overlay network and the physical underlying one. This survey is on the recent research developments in alleviating the problem, in unstructured schemes, using location-aware overlay construction algorithms.
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}
\emph{Peer-to-peer} systems gained significant attention due to their plethora of unique features in supporting file sharing among huge numbers of networked users, where each such \emph{peer} acts both as a resource provider and a resource consumer. Such serverless systems, proved to be able to achieve outstanding aggregate resource capacities as participants join the system\footnote{In peer-to-peer file-sharing systems there has been observed mitigation of this self-scaling property by the \emph{free-rider} problem\cite{adar_gnutellafreeriders_2000}}, without requiring additional expenditure for infrastructure.

As these systems evolved, three main architectures have been emerged:
\begin{itemize}
  \item \emph{centralized}
  \item \emph{decentralized structured}
  \item \emph{decentralized unstructured}
\end{itemize}

\emph{Centralized} architectures were the first to recognize that requests for popular content need not be sent to a central server but instead could be handled by the many hosts that already posses the content. \emph{Napster}, for example, maintained a \emph{central index server} based on file lists provided by participating peers. The central index server was queried by the users and it returned pointers to the actual content. Thus, by centralizing search while distributing downloads, Napster achieved a high functional design that was widely acknowledged, at the time, as ``the fastest growing Internet application ever''. Of course, the architecture is not without flaws. The centralized search facility is its ``Achilles' heel'' as it can be prooven a bottleneck for the system's scalability as well as a central point of failure and vulnerable to \emph{Denial-of-Service (DoS)} attacks.

This centralized approach, has been eventually, replaced by \emph{decentralized}, that distributed both the search and the download capabilities. In the \emph{decentralized unstructured} architectures, file placement is random, which means there is no correlation with the network topology whatsoever \cite{yang_improvep2psearch_2002}. The most important properties of such systems are that they:
\begin{itemize}
  \item support inherent heterogeneity of peers,
  \item are highly resilient to peer failures, and
  \item incur low maintenance overhead at peer arrivals and departures.
\end{itemize}
In the literature, they can be also found under the term \emph{broadcast-based} systems, because they use \emph{message flooding} among peers (Gnutella\cite{gnutella}) or among superpeers (KaZaA\cite{kazaa}) to propagate queries but with no specific destination. Each (super)peer receiving a query
\begin{inparaenum}[\itshape i\upshape)]
  \item decreases the incoming query message's lifetime which is measured in \emph{time-to-live (TTL)} hops,
  \item back-propagates its results, and
  \item if the lifetime of the incoming message did not expired (TTL did not reach zero), it forwards the message to all its neighbours except from the incoming link.
\end{inparaenum}
Early incarnations of the Gnutella protocol, though were considered far from scalable \cite{ritter_gnucantscale_2001} because:
\begin{itemize}
  \item randomly choosing logical neighbours without knowledge of the underlying network causes mismatch between physical and application-level topologies, and/or
  \item queries may be flooded to multiple paths merging in one peer while in such case one of the paths would have been enough, and/or
  \item peers may forward the same message to each other before they receive the query messages from each other
\end{itemize}
More generally, studies upon peer-to-peer traffic \cite{seroiu_analysiscds_2002, sen_analyzep2ptraffic_2004} with measurements on some popular systems, such as FastTrack, Gnutella and DirectConnect have shown that it contributes the largest portion of the overall Internet traffic. \cite{matei_mapgnutella_2002} showed that, even given that 95 percent of any two nodes are less than 7 hops away from each other, a flooding-based algorithm, like the one used by Gnutella, can generate 330TB/month in a 50.000 node network!

Recently \emph{decentralized structured} schemes have been proposed \cite{ratnasamy_can_2001, stoica_chord_2001, zhao_tapestry_2001, antony_pastry_2001}. These, form an application-level topology that provides a hash table look-up and routing service, called \emph{Distributed Hash Table (DHT)} that requires only $O \left ( log n \right )$ steps for a hash-table-like \texttt{look-up()} search that also guarantees perfect recall even if only a single copy of the searched data exists in the system\footnote{In comparison to Gnutella, that requires $O \left ( n \right )$ to reliably locate a specific file}.

Studying the behaviour of the various peer-to-peer schemes \cite{matei_mapgnutella_2002, lv_randomwalks_2002, merugu_str2unstr_2003} showed that the ad-hoc network topology of unstructured overlay networks that preserve \emph{Power Law} and \emph{Small World} characteristics\footnote{Power Law describes the node degree while the Small World describes characteristics of path length and clustering coefficient. The clustering coefficient for a node $\upsilon$ in a graph $G = \left( V, E \right)$ is defined as the ratio of the existing connections between $\upsilon$'s neighbouring nodes to $\gamma \times \left( \gamma - 1 \right)$, where $\gamma$ is the number of neighbouring nodes of $\upsilon$. High cluster coefficient means that neighbouring nodes of any node $\upsilon$ likely connect one another.} \cite{faloutsos_powerlaw_1999, saroiu_measurefileshare_2002} offer a more promising approach. Particularly:
\begin{itemize}
  \item Peer-to-peer clients are extremely \emph{transient}. Unstructured systems can have high maintenance traffic in delivering messages, updating the mapping, discovering failures and replicating lost data or pointers, making them insufficient on highly volatile networks.
  \item \emph{Keyword searches} versus \emph{exact-match queries}. In DHTs there is a tight control between the data placement and the topology of the network. For this reason it is hard to efficiently support partially matched queries while Gnutella and other similar systems effortlessly support keyword searches and other complex queries since the mechanism is realized locally, on a node-by-node basis.
  \item Popular content is located at multiple peers and thus it is more likely for a flooding-based search to return results. DHTs, on the other hand, fit better in the systems which require ability to reliably locate content, even in the extreme case that only a single-copy exists in the network.
\end{itemize}
For this reasons, efforts have been placed for optimizing the efficiency of decentralized unstructured peer-to-peer networks. Research mainly focuses on
\begin{inparaenum}[\itshape i\upshape)]
  \item reducing unnecessary, redundant communication traffic, and
  \item exploiting physical locality to reduce communication response.
\end{inparaenum}
The goal can be achieved at, both, the application-level network as well as the underlying physical one. In the first case by refining the message relay techniques, while in the second one, by adaptively reconstructing the application network to map as well as possible to the the physical network.

This survey begins with a quick review of the conception of the overlay network. It continues with the phenomenon of the topology mismatch between the overlay and the underlying networks and ultimately end up presenting a series of distinguished research work that has been published in the recent literature concerning the alleviation of such a problematic state, focusing on decentralized unstructured peer-to-peer architectures.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Overlay Networks}
An \emph{overlay} network is an abstract, logical interconnection formed on top of another network. It consists of a set of nodes, the computing elements, that are being connected by virtual (logical) point-to-point links. For example, the Asymmetric Digital Subscriber Line (ADSL) is, actually, an overlay network on top of the Public Switched Telephone Network (PSTN).

The abstraction of overlay networks has been proposed as a way to implement efficient, fully distributed, application layer, services on top of the best-effort IP layer forming a widely ranged family of protocols collectively known as \emph{peer-to-peer} or \emph{p2p} for short.

Initial efforts targeted on the implementation of one-to-many addressing schemes that could replace \emph{IP multicast} in providing higher quality of streaming media through \emph{quality of service (QoS)} guarantees. IP multicast as well as other proposals such as \emph{IntServ} and \emph{DiffServ}\cite{cisco_diffserv_2005} have not seen wide acceptance (yet?), largely because they require changes to all routers of the network. On the other hand, an overlay network can be incrementally deployed on end-systems running the overlay protocol software, without cooperation from the ISPs. Academic research includes \cite{chu_esm_2000}, \cite{jannotti_overcast_2000} and \cite{kwon_tag_2002} among others.

Overlay networks were also implemented in order to back the routing of messages to destinations whose address is not known in advance. This resulted in the appearance of well kwown peer-to-peer protocols such as \cite{gnutella} and \cite{maymounkov_kademlia_2002} widely used for digital content sharing among their network nodes.

Other special perpose overlay networks include \cite{anderson_ron_2001} for resilient routing; \cite{subramanian_overqos_2004} for quality of service guarantees; and \cite{clarke_freenet_2001} for anonymity; to name just a few.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The Topology Mismatch Problem}
An overlay network is able to control the sequence of peers, a message traverses, before reaching its destination but it is completely unaware of how the actual packets are switched at the underlying infrastructure along the overlay path. For example, a single logical point-to-point link on the overlay, most of the time corresponds to multiple physical links, in the underlying layer.

Additionally a link in the underlying network often serves the mapping of several overlay paths causing increase of the physical link's \emph{stress}\cite{chu_esm_2002}.  Furthermore, the stochastic behaviour of peers randomly joining and leaving the peer-to-peer network all cause the overlay to physical mapping to create such unnecessary, redundant traffic that impacts the efficiency of the network in terms of average response time. The phenomenon is reffered to as the \emph{topology mismatch problem}.

@@

@@

Several protocols for dealing with the problems derived from the topology mismatch problem in decentralised unstructured peer-to-peer systems have been proposed in the literature. Each implementation has different goals but all share some common ground on the aspect of the network functionality they focus on, in order to achieve them. Thus, algorithms can be categorized into three basic types:
\begin{itemize}
  \item \emph{forwarding-based}. A peer selects only a subset of its neighbours to re-broadcast query messages. The selection is made using one or a combination of various statistical metrics. Examples of such metrics, could be the number of responses received by a neighbour or the connection latency of the link between the nodes e.t.c. These approaches enhance search efficiency but have also several drawbacks. First they drastically reduce the search scope. Expanding the search scope, on the other hand, is no easy task because the overhead of forming multicast trees is proportional to the multicast group size. Also such schemes do not consider dynamic joining and leaving of peers so they do not scale well on dynamic environments.
  \item \emph{cache based}. These approaches can be further divided into:
  \begin{itemize}
    \item \emph{data index caching}, and
    \item \emph{content index caching}.
  \end{itemize}
  Additionally caching policies can be divided into:
  \begin{itemize}
    \item \emph{centralized}, and
    \item \emph{local}.
  \end{itemize}
  Centralized peer-to-peer systems provide centralized index servers to keep indices of shared files of all peer (e.g. Napster). In the local caching policy, each or some peers maintain information about other peers (like the superpeer scheme of Kazaa). Caching can significantly reduce traffic costs and response time but duplication of messages still exist, limiting scalability of such approaches.
  \item \emph{overlay optimization}. This approach can be subdivided furtherlly into:
  \begin{itemize}
    \item \emph{Spanning tree based}. Construction of a rich graph and the build its minimum spanning tree. These approaches incur large traffic overhead to the system \cite{chu_esm_2000, chu_esm_2002}.
    \item \emph{Cluster based}. Selection to link physically closer nodes with each other. These approaches can shrink the search scope significantly while mapping accuracy is not always guaranteed.
    \item \emph{Minimum latency first}. Use of latency as a metric to calculate distance among peers. They require global latency information ``landmarks''\footnote{Measuring latency between peers and stable Internet servers.}.
  \end{itemize}
\end{itemize}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\subsection{Narada}
\emph{End System Multicast} \cite{chu_esm_2000, chu_esm_2001, chu_esm_2002} as well as other  proposes an architecture, where end-systems implement all multicast related functionality (including membership management and packet replication). In \emph{Narada}, the protocol developed in the paper's context, end-systems
\begin{inparaenum}[\itshape i\upshape)]
  \item self-organize, in a fully distributed manner,
  \item into an efficient overlay structure, both from the network and the application perspective,
  \item with self improving properties,
  \item adapting to network dynamics\footnote{Long-term variations in Internet path characteristics (i.e. bandwidth, latency e.t.c.)}.
\end{inparaenum}

What Narada really does is to construct overlay spanning trees in a two-step process:
\begin{enumerate}
  \item It builds a richer connected graph \footnote{ENarada \cite{li_enarada_2008} used Gossip protocol for the construction}, termed a \emph{mesh}, while trying to ensure a couple of desired performance properties:
    \begin{itemize}
      \item path quality between any pair of members should be comparable to the quality of the unicast path between them, and
      \item each member should have a limited number of neighbours on the mesh.
    \end{itemize}

Narada incrementally improves the quality of the mesh by adding and dropping overlay links. Members probe each other at random and new links are added depending on what is the gain in doing so. On the other hand, participants, continuously  monitor the utility of existing links, and drop them if they are found to be not usefull.

Narada protocol achieves relatively good performance for small and medium sized groups involving tens to hundreds of members. This is not, by no means, what a contemporary peer-to-peer system needs to be characterized as scalable, but this work is considered pioneering\footnote{Another similar approach is \emph{Scattercast}\cite{chawathe_scattercast_2000}}, in that it was the first to have conducted a detailed Internet evaluation to analyse and demonstrate the feasibility of an overlay-based, application layer, multicast architecture that can dynamically adapt to bandwidth and latency properties of the physical underlying infrastructure.

\subsection{Gia}
\emph{Gia} \cite{chawathe_gia_2003} is a decentralized unstructured peer-to-peer file sharing system with the following characteristics:
\begin{inparaenum}[\itshape i\upshape)]
  \item replaces Gnutella's flooding with random walks \cite{lv_randomwalks_2002},
  \item recognizes the importance of overlay network's topology, while using the random walks, and exploits it by incorporating a topology adaptation algorithm,
  \item introduces a token-based flow control mechanism, and
  \item detects the significant heterogeneity in peer bandwidth, processing power, disk speed e.t.c. and takes advantage of it.
\end{inparaenum}

More specifically, there are four key components in the design of \emph{Gia} which are summarized bellow:
\begin{enumerate}
  \item A \emph{dynamic topology adaptation} protocol that puts participating nodes within short reach of high capacity nodes so that these \emph{high-degree} nodes, which due to their high connectivity receive most of the queries, actually have the capacity to handle them.
  \item An \emph{active flow control} scheme is used to avoid overloaded hot-spots. Heterogeneity is detected and flow control tokens are given to nodes based on the available capacity.
  \item Every node maintains pointers to to the content that is offered by their immediate neighbours, creating a \emph{one-hop replication} of pointers, scheme.
  \item A \emph{search protocol} based on random walks, that is biased towards directing queries to high-capacity nodes that are typically best able to answer these queries.
\end{enumerate}

\subsection{Adaptive Overlay Topology Optimization}
\emph{Adaptive Overlay Topology Optimization (AOTO)} \cite{liu_auto_2003} is an algorithm for building an overlay multicast tree among each source node and its direct logical neighbours in order to alleviate the mismatching problem while providing a larger query coverage range. It includes the steps of
\begin{inparaenum}[\itshape i\upshape)]
  \item \emph{Selective Flooding (SF)}, and
  \item \emph{Active Topology (AT)}
\end{inparaenum}
which are furtherlly discussed in the following paragraphs.

\subsubsection*{Selective Flooding (SF)}
Instead of flooding to all neighbours, SF builds an overlay multicast tree among each peer and its immediate logical neighbours. This reduces overall message burden to the network by making some neighbours non-flooding. The tree is formed using a minimum spanning tree algorithm. For this reason a peer needs to know the costs to all its logical neighbours (e.g. the network delay) as well as between any pair of neighbours. Additional, cost probing, messages need to be added to the original messages defined by the Gnutella protocol.

\subsubsection*{Active Topology (AT)}
At this step the overlay topology is reorganized. Each peer, independently, makes optimizations in the overlay network to alleviate topology mismatching by replacing non-flooding neighbours, with closer nodes. The approach incorporated by the paper is called \emph{Randomized AT} algorithm which picks up a candidate peer at random among the non-flooding neighbour's neighbours. Whenever a new neighbour cost table is received or there is a change of neighbours, the source peer has to re-calculate the multicast tree and apply the randomized AT algorithm.

\subsection{Location-aware Topology Matching}
\emph{Location-aware topology matching} or \emph{LTM} \cite{liu_ltm_2004}, is an algorithm of building an efficient overlay by disconnecting low productive connections and choosing physically closer nodes as logical neighbours while still retaining the search scope and reducing response time of queries. This is done by issuing detector messages by peers, to a small region so that receivers of the detector can record relative delay information and cut inefficient or redundant logical links as well as add closer nodes as direct neighbours.

Three operations are defined in LTMband are furtherly discussed in the following paragraphs.

\subsubsection*{TTL2-detector flooding}
\emph{TTL2-detector} messages are designed based on the specification of the Gnutella protocol. Suppose that $d(i, S, v)$ denotes the TTL2-detector which has the message ID of $i$ with TTL value of $v$ and is initiated by $S$. In addition to Gnutella's unified 23-byte header, such a message has a body of two types:
\begin{itemize}
  \item short format $d(i, S, 1)$ if $S$ is a direct neighbour of the receiver
  \item long format $d(i, S, 0)$ if $S$ is within a two-hop distance from the receiver
\end{itemize}
This way a peer can calculate link costs from within a one- or two-hop area around it and thus reach peers in $N(S)$ and $N^2(S)$ sets, respectively, in the paper's context. Each peer floods a TTL-detector periodically while it calculates link cost to another peer by receiving other peer's detector messages and checking:
\begin{itemize}
  \item the Source Timestamp field and its own clock upon time of receipt for one-hop messages
  \item the Source Timestamp, TTL1 Timestamp and its own clock upon receipt for two-hop messages
\end{itemize}
Of course peer clocks must be synchronized in order to do such checking.

\subsubsection*{Low productive connection cutting}
There are three cases for any peer $P$ who receives $d(i, S, v)$ multiple times:
\begin{inparaenum}[\itshape i\upshape)]
  \item $P$ receives both $d(i, S, 1)$ and $d(i, S, 0)$
  \item $P$ receives multiple $d(i, S, 0)$s from different paths, and P randomly chooses to process one
  \item $P$ receives one $d(i, S, 1)$ and multiple $d(i, S, 0)$s, and $P$ processes $d(i, S, 1)$ and one randomly selected $d(i, S, 0)$
\end{inparaenum}
If the link with the largest cost is found and is a direct neighbour then the connection is put in a will-cut list and stays there for a certain period of time. If it is not, then it is handled by other peers. After that period, connections are cut and recorded to $P$'s cut-list.

\subsubsection*{Source probing}
For a peer $P\in(N^2(S) - N(S))$ who receives only one $d(i, S, 0)$, the cost of $PS$ is obtained (with list look-up or probing). Then $P$ compares it with the cost form each hop and if $PS$ has the largest cost, $P$ will not keep this connection, while otherwise the connection will be created.

LTM disadvantages are
\begin{inparaenum}[\itshape i\upshape)]
  \item disagreement of measured delay due to unsynchronized clocks causes problems when deciding the cut positions, which can influence the network connectivity, and
  \item the network delay metric mainly focuses on disabling the connections between peers physically far away without considering the shortcuts created by powerful peers.
\end{inparaenum}

\subsection{Adaptive Connection Establishment}
\emph{Adaptive Connection Establishment (ACE)} \cite{liu_ace_2004} builds an overlay multicast tree among each source node and the peers within a certain diameter from the source peer and optimizes the neighbour connections that are not in that tree.

>>>>

%opening
% TODO: Contribution of this paper
% Q&A: 'Overlay' or 'Peer-to-Peer'/'P2P'
\title{Alleviating the Topology Mismatch Problem in Unstructured and Structured Overlay Networks: A Survey}
\author{
Vassilis Moustakas\\
Department of Informatics and Telecommunications,\\
National and Kapodistrian University of Athens\\
\mbox{\texttt{b.moystakas@di.uoa.gr}}
}

\begin{document}
\maketitle

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\begin{abstract}
% TODO: Contribution of this paper
Peer-to-peer (p2p) systems is a rapidly growing application architecture. The overlay network abstraction on top of a best-effort infrastructure provides a powerfull tool to realize ever-wanted exotic features like anonymity, high availability and robustness, load balancing, QoS at the application layer and many more. Unfortunately weaknesses of early effectuations constrained applications from unleashing the full potential of the paradigm. One of them is the topology mismatch problem between the overlay and the physical underlying networks, a problem that imposes unecessary stress to the network resources. This survey investigates on the recent research developments towards the alleviation of the problem, in both unstructured and structures p2p schemes.
\end{abstract}

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Introduction}

% TODO: Contribution of this survey
% UNDER CONSTRUCTION

% TODO: Structure of this paper
% UNDER CONSTRUCTION!
This survey begins with a quick review of the conception of the overlay network and the available architectures of p2p systems. It continues with the identification of the phenomenon of the topology mismatch between overlay and underlying networks and its impact on the utilization of network resources. Ultimately, it reviews the most important resourses found in the literature that attempt to tackle the problem as it proposes a taxonomic scheme, using a set of supertype-subtype relationships based on unique characteristics they have and/or specific goals they target.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% OK
\section{Overlay Networks and P2P Systems}
An \emph{overlay network} is an abstract, logical interconnection of entities formed on top of another network. It consists of a set of nodes, the computing elements, that are being connected by virtual (logical) point-to-point links. For example, the \emph{Asymmetric Digital Subscriber Line (ADSL)} is, actually, an overlay network on top of the \emph{Public Switched Telephone Network (PSTN)}.

The abstraction of overlay networks has been proposed as a way to implement efficient, fully distributed, application layer services on top a best-effort IP layer forming a widely ranged family of protocols collectively known as \emph{peer-to-peer} or \emph{p2p} for short. P2p network systems gained significant attention, in recent years, due to their plethora of unique features in supporting file sharing among huge numbers of (inter-) networked users, where each, such, \emph{peer} could act both as a resource provider and a resource consumer. This lead to the introduction of the \emph{servent}-concept \cite{gnutella}, a portmanteau that blends the notions of \emph{server} and \emph{client} to denote the twofold role of the participants in a p2p network. Such serverless systems, proved to be able to achieve outstanding aggregate resource capacities as participants join the system\footnote{Unfortunatelly, there has been observed mitigation of this self-scaling property by the undesirable behaviour of participating users, usually answered in popular file-sharing systems; from \emph{free-riding} (i.e. in Napster and Gnutella networks \cite{saroiu_measurefileshare_2002, adar_gnutellafreeriders_2000, hughes_gnutellafreeride_2005}), to the distribution of illegal content and/or other \emph{socio-technical} issues \cite{hughes_socp2p_2008}.}, without requiring additional expenditure for infrastructure.

As p2p systems evolved, three main architectures have been emerged, namely:
\begin{itemize}
  \item \emph{centralized}
  \item \emph{decentralized structured}
  \item \emph{decentralized unstructured}
\end{itemize}

\emph{Centralized} architectures were the first to recognize that requests for popular content need not be sent to a central server but instead could be handled by the many hosts that already posses the content. \emph{Napster}, for example, maintained a \emph{central index server} based on file lists provided by participating peers. The central index server was queried by the users and it returned pointers to the actual content. Thus, by centralizing search while distributing downloads, Napster achieved a highly functional design that was widely acknowledged, at the time, as ``the fastest growing Internet application ever''. This centralized search facility, though, renders this scheme not fully distributed and can be proven its ``Achilles' heel'' in terms of being a scalability bottleneck, a central point of failure and vulnerable to malicious acts (e.g. \emph{Denial-of-Service (DoS) attacks}).

This centralized approach, has been eventually, replaced by architectures that distributed both the search and the download capabilities. In these \emph{decentralized unstructured} architectures, file placement is random, which means there is no correlation with the network topology whatsoever \cite{yang_improvep2psearch_2002}. The most important properties of such systems are that they support inherent heterogeneity of peers, are highly resilient to peer failures, and incur low maintenance overhead at handling the dynamics of peer participation \cite{stutzbach_churn_2006}. In the literature, they are also known as \emph{broadcast-based} systems, because they use \emph{message flooding} among peers (i.e. Gnutella\cite{gnutella}) or among super-peers (i.e KaZaA\cite{kazaa}) to propagate queries. Each (super-) peer receiving a query decreases the incoming query message's lifetime which is measured in \emph{time-to-live (TTL)} hops, back-propagates its results, and  if the lifetime of the incoming message did not expired (TTL did not reach zero), it forwards the message to all its neighbours except from the one it receives it from.
% TODO: maybe enumerate applications of decentralized unstructured p2ps.

Recently, \emph{decentralized structured} schemes have been proposed in order to provide a self-organising infrastructure for large-scale p2p applications \cite{ratnasamy_can_2001, stoica_chord_2001, antony_pastry_2001, zhao_tapestry_2001, maymounkov_kademlia_2002}. They implement a \emph{Distributed Hash Table (DHT)} that maps objects to nodes through a deterministic mechanism. The main advantage of the DHT approach of decentralized structured p2p networks over decentralized unstructured ones, is that they provide a guaranteed bound on the number of overlay routing hops that have to be taken in order for an object to be located even in the case where only a single copy exists in the system. This is $O \left ( log n \right )$ compared to Gnutella, that requires $O \left ( n \right )$ to reliably locate a specific object.
% TODO: maybe enumerate applications of decentralized structured p2ps. (OceanStore, Far)

% TODO: enumerate applications of overlay networks and p2p systems regardless of architecture
% UNDER CONSTRUCTION
% Q&A: Need more??
Initial efforts targeted on the implementation of one-to-many addressing schemes that could replace \emph{IP multicast} in providing higher quality of streaming media through \emph{quality of service (QoS)} guarantees. IP multicast as well as other proposals such as \emph{IntServ} and \emph{DiffServ} \cite{cisco_diffserv_2005} have not seen wide acceptance (yet?), largely because they require changes to all routers of the network. On the other hand, an overlay network can be incrementally deployed on end-systems running the overlay protocol software, without cooperation from the ISPs. Academic research includes \cite{chu_esm_2000, jannotti_overcast_2000, kwon_tag_2002} among others.
Overlay networks were also implemented in order to back the routing of messages to destinations whose address is not known in advance. This resulted in the appearance of well kwown peer-to-peer protocols such as \cite{gnutella} and \cite{maymounkov_kademlia_2002} widely used for digital content sharing among their network nodes.
Other special perpose overlay networks include \cite{anderson_ron_2001} for resilient routing; \cite{subramanian_overqos_2004} for quality of service guarantees; and \cite{clarke_freenet_2001} for anonymity; to name just a few.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% start
%TODO: Maybe this section should be reviewed or deleted or go to another section
%Studying the behaviour of the various peer-to-peer schemes \cite{matei_mapgnutella_2002, lv_randomwalks_2002, merugu_str2unstr_2003} showed that the ad-hoc network topology of unstructured overlay networks that preserve \emph{Power Law} and \emph{Small World} characteristics\footnote{Power Law describes the node degree while the Small World describes characteristics of path length and clustering coefficient. The clustering coefficient for a node $\upsilon$ in a graph $G = \left( V, E \right)$ is defined as the ratio of the existing connections between $\upsilon$'s neighbouring nodes to $\gamma \times \left( \gamma - 1 \right)$, where $\gamma$ is the number of neighbouring nodes of $\upsilon$. High cluster coefficient means that neighbouring nodes of any node $\upsilon$ likely connect one another.} \cite{faloutsos_powerlaw_1999, saroiu_measurefileshare_2002} offer a more promising approach. Particularly:
%\begin{itemize}
%  \item Peer-to-peer clients are extremely \emph{transient}. Unstructured systems can have high maintenance traffic in delivering messages, updating the mapping, discovering failures and replicating lost data or pointers, making them insufficient on highly volatile networks.
%  \item \emph{Keyword searches} versus \emph{exact-match queries}. In DHTs there is a tight control between the data placement and the topology of the network. For this reason it is hard to efficiently support partially matched queries while Gnutella and other similar systems effortlessly support keyword searches and other complex queries since the mechanism is realized locally, on a node-by-node basis.
%  \item Popular content is located at multiple peers and thus it is more likely for a flooding-based search to return results. DHTs, on the other hand, fit better in the systems which require ability to reliably locate content, even in the extreme case that only a single-copy exists in the network.
%\end{itemize}
% end
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

% TODO: how 
%For this reasons, efforts have been placed for optimizing the efficiency of decentralized unstructured peer-to-peer networks. Research mainly focuses on
%\begin{inparaenum}[\itshape i\upshape)]
%  \item reducing unnecessary, redundant communication traffic, and
%  \item exploiting physical locality to reduce communication response.
%\end{inparaenum}
%The goal can be achieved at, both, the application-level network as well as the underlying physical one. In the first case by refining the message relay techniques, while in the second one, by adaptively reconstructing the application network to map as well as possible to the the physical network.

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{The Topology Mismatch Problem}

% TODO: Define the problem
% UNDER CONSTRUCTION
% Q&A is it too late??
An overlay network is able to control the sequence of peers, a message traverses, before reaching its destination but it is completely unaware of how the actual packets are switched at the underlying infrastructure along the overlay path. For example, a single logical point-to-point link on the overlay, most of the time, corresponds to multiple physical links, in the underlying layer. Additionally a link in the underlying network often serves the mapping of several overlay paths causing increase of the physical link's \emph{stress}\cite{chu_esm_2002}.  Furthermore, the stochastic behaviour of peers randomly joining and leaving the peer-to-peer network all cause the overlay to physical mapping to create such unnecessary, redundant traffic that impacts the efficiency of the network in terms of average response time. The phenomenon is reffered to as the \emph{topology mismatch problem}.

@@

% TODO: Argue on the problem
% UNDER CONSTRUCTION
Early incarnations of the Gnutella protocol, though were considered far from scalable \cite{ritter_gnucantscale_2001} because randomly choosing logical neighbours without knowledge of the underlying network causes mismatch between physical and application-level topologies. Additionaly, queries may be flooded to multiple paths merging in one peer while in such case one of the paths would have been enough. Nevertheless, peers may forward the same message to each other before they receive the query messages from each other.

In structured schemes, typically, node IDs are assigned \emph{randomly}, resulting in excellent load balancing, scalability and robustness. Unfortunatelly, this randomness has a negative impact on the \emph{routing locality} of the network. This means that even though the target node can be reached with logaritmic overlay hops, the distance traveled in the physical underlying network, during the overlay routing process, can be far from optimal. The ratio of the phyical distance, a query travels through the overlay to an object, and the minimal distance to that object (i.e. though IP) is known in the literature as \emph{stretch} or \emph{Relative Delay Penalty (RDP)}.

Studies on peer-to-peer traffic \cite{seroiu_analysiscds_2002, sen_analyzep2ptraffic_2004} with measurements on some popular systems, such as FastTrack, Gnutella and DirectConnect have shown that it contributes the largest portion of the overall Internet traffic. \cite{matei_mapgnutella_2002} showed that, even given that 95\% of any two nodes are less than 7 hops away from each other, a flooding-based algorithm, like the one used by Gnutella, can generate 330TB/month in a 50.000 node network!

@@

% TODO: Taxomize the approaches that have been proposed in the literature
% for unstructured and structured overlay networks
% UNDER CONSTRUCTION
% VERY IMPORTANT SECTION
\section{Taxonomy}

% TODO: Unstructured
% UNDER CONSTRUCTION
Several protocols for dealing with the problems derived from the topology mismatch problem in decentralised unstructured peer-to-peer systems have been proposed in the literature. Each implementation has different goals but all share some common ground on the aspect of the network functionality they focus on, in order to achieve them. Thus, algorithms can be categorized into three basic types:
\begin{itemize}
  \item \emph{forwarding-based}. A peer selects only a subset of its neighbours to re-broadcast query messages. The selection is made using one or a combination of various statistical metrics. Examples of such metrics, could be the number of responses received by a neighbour or the connection latency of the link between the nodes e.t.c. These approaches enhance search efficiency but have also several drawbacks. First they drastically reduce the search scope. Expanding the search scope, on the other hand, is no easy task because the overhead of forming multicast trees is proportional to the multicast group size. Also such schemes do not consider dynamic joining and leaving of peers so they do not scale well on dynamic environments.
  \item \emph{cache based}. These approaches can be further divided into:
  \begin{itemize}
    \item \emph{data index caching}, and
    \item \emph{content index caching}.
  \end{itemize}
  Additionally caching policies can be divided into:
  \begin{itemize}
    \item \emph{centralized}, and
    \item \emph{local}.
  \end{itemize}
  Centralized peer-to-peer systems provide centralized index servers to keep indices of shared files of all peer (e.g. Napster). In the local caching policy, each or some peers maintain information about other peers (like the superpeer scheme of Kazaa). Caching can significantly reduce traffic costs and response time but duplication of messages still exist, limiting scalability of such approaches.
  \item \emph{overlay optimization}. This approach can be subdivided furtherlly into:
  \begin{itemize}
    \item \emph{Spanning tree based}. Construction of a rich graph and the build its minimum spanning tree. These approaches incur large traffic overhead to the system \cite{chu_esm_2000, chu_esm_2002}.
    \item \emph{Cluster based}. Selection to link physically closer nodes with each other. These approaches can shrink the search scope significantly while mapping accuracy is not always guaranteed.
    \item \emph{Minimum latency first}. Use of latency as a metric to calculate distance among peers. They require global latency information ``landmarks''\footnote{Measuring latency between peers and stable Internet servers.}.
  \end{itemize}
\end{itemize}

% TODO: Structured
% UNDER CONSTRUCTION
Xu et.al in \cite{xu_globstate_2003} state that there are three ways generaring proximity information
\begin{itemize}
 \item Expanding Ring Search,
 \item Heuristics, and
 \item Landmark Binning
\end{itemize}
